---
title: "Marketing ROI Analysis"
author: "Aryaman Dhindsa"
date: April 2,2026
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: flatly
    self-contained: true # bundles everything into one html file
  pdf:
    number-sections: true
    fontsize: 11pt
---



```{r Loading Libraries}
#| echo: false
#| warning: false
#| message: false
# Loading all the required libraries
library(tidyverse)
library(knitr)
library(lubridate)
```

# Introduction
Over the past quarter, we implemented a personalised marketing campaign where spending varied across customers through ads, emails, and promotional offers. This approach aimed to better target individuals based on expected value and engagement. The campaign data provides an opportunity to evaluate how effective this experimental variation in marketing investment was.

# Research Question
Is there a positive or a negative relationship between marketing spend and customer purchase amounts across different customer segments (e.g., Enterprise and Consumer)?

# Dataset Description

This dataset records customer-level marketing return on investment (ROI) information. Each observation represents a unique customer, identified by a costumer ID, and tracks the date they signed up, the amount spent on marketing to engage them, and their subsequent purchase amount. Customers are also categorised into one of the market segments: Consumer, SMB, or Enterprise. Together, these variables allow for analysis of marketing effectiveness and revenue generation across different customer groups.

## Cleaning process

```{r loading and cleaning data}
#| echo: true
#| include: false

# loading the dataset
roi_data <- read.csv("data/35683422_Aryaman Dhindsa Dhindsa.csv")

# Now, cleaning the dataset 
clean_roi_data <- roi_data |> 
    select(-notes, -visits) |>
    mutate(
    signup_date = na_if(signup_date, "N/A"),
    signup_date = na_if(signup_date, "NULL"),
    signup_date = parse_date_time(signup_date, 
                  orders = c("dmy", "ymd", "mdy"), 
                  quiet = TRUE),
    signup_date = as.Date(signup_date)
    ) |> 
    drop_na()
```

Two columns — `notes` and `visits` — were removed as they were not relevant to the analysis. The `signup_date` column contained inconsistently formatted entries as well as `"N/A"` and `"NULL"` strings, which were converted to true `NA` values before being parsed across multiple date formats (day-month-year, year-month-day, and month-day-year) and coerced to a uniform `Date` type. Rows with any remaining missing values were then removed via `drop_na()`. Thus, making the dataset clean.

## Dataset dimensions

The cleaned dataset contains `r nrow(clean_roi_data)` observations across `r ncol(clean_roi_data)` variables: `customer_id` (str), `signup_date` (Date), `marketing_spend` (dbl), `purchase_amount` (dbl), and `segment` (str).

